---
title: "Vectors and Matrices"
author: Lan Huong Nguyen
tutorial:
  id: "com.exercises.vectors-matrices"
  version: 0.1
output: 
  learnr::tutorial:
     progressive: true
runtime: shiny_prerendered
---

options(tutorial.exercise.timelimit = 10)

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = checkr::checkr_tutor)
#tutorial_options(exercise.checker = checkthat::check_exercise)
```


## Vector Basics

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*

### Exercise 1

Using the function seq(), write the R code to create a vector named `myvector`,
which is a sequence of numbers from 15 to 85 at increments of 5, i.e. a sequence
15, 20, 25, ..., 75, 80, 85:

```{r pS1}
myvector <- seq(15, 85, by = 5)
```


```{r s1e1, exercise=TRUE}
myvector <- 
```


```{r s1e1-hint}
# To see the documentation of seq() and see how the function works you can run ?seq.
myvector <- seq(..., ..., by = ...)
```

```{r s1e1-solution}
myvector <- seq(15, 85, by = 5)
```


```{r s1e1-code-check}
check_blanks(USER_CODE)
```

```{r s1e1-check}
test_0 <- find_call("seq(whatever)", "you need to use seq() function")
test_1 <- find_call("seq(whatever, whatever, whatever)", "you need to provide 3 arguments to seq(): from, to, by")
test_2 <- find_assignment("myvector", "you need to create a vector named 'myvector'")
test_3 <- check_value(match_vector(seq(15, 85, by = 5), hint = TRUE), "values do not agree")

USER_CODE %>% test_0 %>% test_1 %>% test_2 %>% final_ %>% test_3
```


### Exercise 2

Divide each element of 'myvector' by 5.

```{r s1e2, exercise=TRUE, exercise.setup = "pS1"}

```

```{r s1e2-code-check, echo = FALSE}
check_blanks(USER_CODE)
```


```{r s1e2-solution}
myvector / 5
```

```{r s1e2-check}
test_1 <- find_call("whatever / whatever", "you need to use division (/)")
test_2 <- check_value(match_vector(seq(15, 85, by = 5)/5), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


### Exercise 3

Raise each element of 'myvector' to the 3rd power.

```{r s1e3, exercise=TRUE, exercise.setup = "pS1"}

```


```{r s1e3-solution}
myvector^3
```

```{r s1e3-check}
test_1 <- find_call("whatever ^ whatever", "you need to use power oprator (^)")
test_2 <- check_value(match_vector(seq(15, 85, by = 5)^3), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


### Exercise 4

Compute the natural logarithm of each element of 'myvector'.

```{r s1e4, exercise=TRUE,  exercise.setup = "pS1"}

```


```{r s1e4-solution}
log(myvector)
```

```{r s1e4-check}
test_1 <- find_call("log(whatever)", "you need to use the function (log())")
test_2 <- check_value(match_vector(log(seq(15, 85, by = 5))), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


### Exercise 5

Subract a vector of numbers 1, 2, 3, ..., 15 from 'myvector'.

```{r s1e5, exercise=TRUE,  exercise.setup = "pS1"}

```

```{r s1e5-hint}
# Use ':' or seq() to generate a vector c(1, 2, ..., 15)
```

```{r s1e5-solution}
myvector - 1:15
# or myvector - seq(1, 15, 1)
```

```{r s1e5-check}
test_1 <- find_call("whatever - whatever", "you need to use subtraction (-)")
test_2 <- check_value(match_vector(seq(15, 85, by = 5) - 1:15), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```

### Exercise 6

Subract a vector c(1, 2, 3, 4) from 'myvector'.

```{r s1e6, exercise=TRUE,  exercise.setup = "pS1"}

```


```{r s1e6-solution}
myvector - c(1, 2, 3, 4)
```

```{r s1e6-check}
test_1 <- find_call("whatever - whatever", "you need to use subtraction (-)")
test_2 <- check_value(match_vector(seq(15, 85, by = 5) - c(1, 2, 3, 4)), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


### Question 1


```{r s1q1}
question("What happens when R is trying to add or subtract vectors of different lengths?",
  answer("R returns an error message"),
  answer("R crashes"),
  answer("R recycles values of a shorter vector, and prints out a warning", correct = TRUE),
  answer("R appends arbitrary random numbers to get vectors of equal lengths"),
  answer("R cuts the vectors to the same length"),
  random_answer_order = TRUE,
  allow_retry = TRUE
)
```


### Exercise 7

Print the 1st, 5th and 10th element of 'myvector'

```{r s1e7, exercise=TRUE,  exercise.setup = "pS1"}

```


```{r s1e7-solution}
myvector[c(1, 5, 10)]
```

```{r s1e7-check}
test_1 <- find_statement("myvector\\[(.*)\\]", "you need to use square brackets for indexing", regex = TRUE)
test_2 <- check_value(match_vector(seq(15, 85, by = 5)[c(1, 5, 10)]), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


### Exercise 8

Print the all elements of 'myvector' except for the 3rd and the 14th.

```{r s1e8, exercise=TRUE,  exercise.setup = "pS1"}

```


```{r s1e8-solution}
myvector[c(-3, -14)]
```

```{r s1e8-check}
test_1 <- find_statement("myvector\\[(.*)\\]", "you need to use square brackets for indexing", regex = TRUE)
test_2 <- check_value(match_vector(seq(15, 85, by = 5)[c(-3, -14)]), "values do not agree")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


The end of the section.


## Random numbers and subsets

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*

### Exercise 1

Generate and print a vector of 10 random numbers with values between 1 and 100:

```{r s2e1, exercise=TRUE}

```


```{r s2e1-hint-1}
# Use '1:100' to get a set of numbers from 1 to 100
```

```{r s2e1-hint-2}
sample(...)
```

```{r s2e1-solution}
sample(1:100, size = 10)
```


```{r s2e1-code-check}
check_blanks(USER_CODE)
```

```{r s2e1-check}
test_1 <- find_call("sample(whatever)", "you need to use a sample() function")
test_2 <- find_call("sample(whatever, whatever)", "you provide a size argument")
test_3 <- check_value(agrees(length(x) == 10), "you need to generate 10 values")
test_4 <- check_value(agrees(all(x %in% 1:100)), "the numbers are not between 1 and 100")

USER_CODE %>% test_1 %>% test_2 %>% final_ %>% test_3 %>% final_ %>% test_4 
```

Generate another vector of 10 random numbers with values between 1 and 100:

```{r s2e1i, exercise=TRUE}

```

```{r s2e1i-solution}
sample(1:100, size = 10)
```

```{r s2e1i-check}
test_1 <- find_call("sample(whatever)", "you need to use a sample() function")
test_2 <- find_call("sample(whatever, whatever)", "you provide a size argument")
test_3 <- check_value(agrees(length(x) == 10), "you need to generate 10 values")
test_4 <- check_value(agrees(all(x %in% 1:100)), "the numbers are not between 1 and 100")

USER_CODE %>% test_1 %>% test_2 %>% final_ %>% test_3 %>% final_ %>% test_4 
```

### Exercise 1.2

**Note that:** In the previous section, the two calls to the same function with 
the same input and arguments give two different answers.
This is because the sample() function generates *random* numbers.

Now repeat the same process, but set a random seed 
before generating again the two random vectors.

```{r s2e1ii, exercise=TRUE}
set.seed(123)
sample(1:100, size = 10)
```


```{r s2e1iii, exercise=TRUE}
set.seed(123)
sample(1:100, size = 10)
```


**Note that:** Now two vectors are the same; set.seed() function sets the random number generator.
You can save the seed number and report it or give it to your collaborators, so that they can
reproduce your results if any methods you used involved randomness.

### Exercise 2

First, run the following code to generate the random vector 'rVec' of letters:

```{r prepare-s2e2, echo = TRUE}
set.seed(438567)
rVec <- sample(letters, size = 30, replace = TRUE)
```


```{r s2e2p, exercise=TRUE}
set.seed(438567)
rVec <- sample(letters, size = 30, replace = TRUE)
rVec
```

Now, obtain a random subset of 5 elements of 'rVec'. Note that, this
means you are taking a random subset of a random vector.

```{r s2e2, exercise=TRUE, exercise.setup = "prepare-s2e2"}

```


```{r s2e2-hint}
# Use sample() function
```

```{r s2e2-solution, exercise.setup = "prepare-s2e2"}
sample(rVec, size = 5)
```


```{r s2e2-check, warning=FALSE}
test_1 <- find_call("sample(whatever)", "you need to use a sample() function")
test_2 <- find_call("sample(whatever, whatever)", "you provide a size argument")
test_3 <- check_value(agrees(length(x) == 5), "you need to generate 5 values")
test_4 <- check_value(agrees({set.seed(438567); 
  all(c(x %in% sample(letters, size = 30, replace = TRUE)))}), 
                      "values generated are not in 'rVec'")
USER_CODE %>% test_1 %>% test_2 %>% final_ %>% test_3 
USER_CODE %>% test_1 %>% test_2 %>% final_ %>% test_4 
```

The end of the section.

## Matrices

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*

### Exercise 1

```{r pS3}
X <- matrix(c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18, 
              46, 71, 58, 2, 25, 53, 70, 94, 95, 84, 
              4, 72, 27, 57, 43, 49, 9, 52, 63, 87), 
            nrow = 5, ncol = 6)
```

Create a 5 by 6 matrix named X, with elements equal to 
c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18, 46, 71, 58, 2, 25, 53, 70, 
94, 95, 84, 4, 72, 27, 57, 43, 49, 9, 52, 63, 87) where the elements 
are filled by column:

```{r s3e1, exercise=TRUE}
X <- ...
X
```


```{r s3e1-hint}
# To see the documentation of seq() and see how the function works you can run ?seq.
X <- matrix(c(...), ...)
```

```{r s3e1-solution}
X <- matrix(c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18, 
              46, 71, 58, 2, 25, 53, 70, 94, 95, 84, 
              4, 72, 27, 57, 43, 49, 9, 52, 63, 87), 
            nrow = 5, ncol = 6)
X
```


```{r s3e1-check}
test_1 <- find_call("matrix(whatever)", "you need to use matrix() function")
test_2 <- find_assignment("X", "you need to create a matrix named 'X'")
test_3 <- check_value(agrees(all(dim(x) ==  c(5, 6))), "incorrect matrix dimensions")

USER_CODE %>% test_1 %>% test_2 %>% test_3 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


Create a matrix X2 with the dimensions containing the same elements
but now filled by row

```{r s3e1i, exercise=TRUE}
X2 <- ...
X2
```


```{r s3e1i-hint}
# To see the documentation of seq() and see how the function works you can run ?seq.
X2 <- matrix(c(...), ..., byrow = TRUE)
```

```{r s3e1i-solution}
X2 <- matrix(c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18, 
              46, 71, 58, 2, 25, 53, 70, 94, 95, 84, 
              4, 72, 27, 57, 43, 49, 9, 52, 63, 87), 
            nrow = 5, ncol = 6, byrow = TRUE)
X2
```


```{r s3e1i-check}
test_1 <- find_call("matrix(whatever)", "you need to use matrix() function")
test_2 <- find_assignment("X2", "you need to create a matrix named 'X2'")
test_3 <- check_value(agrees(all(dim(x) ==c(5, 6))), "incorrect matrix dimensions")
USER_CODE %>% test_1 %>% test_2 %>% test_3 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


### Exercise 2

Print a transpose of matrix X. 

```{r s3e2, exercise=TRUE,  exercise.setup = "pS3"}

```



```{r s3e2-hint}
# Use function t() for transposition.
```

```{r s3e2-solution}
t(X)
```


```{r s3e2-check}
test_1 <- find_call("t(whatever)", "you need to use t() function")
USER_CODE %>% test_1
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


### Exercise 3

Rename the column names of X to "F1", "F2", ..., "F6" (features 1 to 6)
and print X.

```{r s3e3, exercise=TRUE,  exercise.setup = "pS3"}

```



```{r s3e3-hint}
# Use function colnames()
colnames(X) <- 
```

```{r s3e3-solution}
colnames(X) <- c("F1", "F2", "F3", "F4", "F5", "F6")
X
```

```{r s3e3-check}
test_1 <- find_call("colnames(whatever)", "you need to use colnames() function")
test_2 <- check_value(agrees(all(colnames(x) ==   c("F1", "F2", "F3", "F4", "F5", "F6"))), 
                      "incorrect column names")
USER_CODE %>% test_1 %>% final_ %>% test_2
```

```{r pS3i}
X <- matrix(c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18,
              46, 71, 58, 2, 25, 53, 70, 94, 95, 84,
              4, 72, 27, 57, 43, 49, 9, 52, 63, 87),
            nrow = 5, ncol = 6)
colnames(X) <- c("F1", "F2", "F3", "F4", "F5", "F6")
```

Rename the row names of X to "S1", "S2", ..."S5" (sample 1 to 5) and print X.

```{r s3e3i, exercise=TRUE,  exercise.setup = "pS3i"}

```



```{r s3e3i-hint}
# Use function rownames() 
rownames(X) <- 
```

```{r s3e3i-solution}
rownames(X) <- c( "S1", "S2", "S3", "S4", "S5")
X
```

```{r s3e3i-check}
test_1 <- find_call("rownames(whatever)", "you need to use rownames() function")
test_2 <- check_value(agrees(all(rownames(x) ==  c( "S1", "S2", "S3", "S4", "S5"))), 
                      "incorrect row names")
USER_CODE %>% test_1 %>% final_ %>% test_2
```


```{r pS3iii}
X <- matrix(c(75, 51, 7, 36, 48, 44, 34, 28, 17, 18,
              46, 71, 58, 2, 25, 53, 70, 94, 95, 84,
              4, 72, 27, 57, 43, 49, 9, 52, 63, 87),
            nrow = 5, ncol = 6)
colnames(X) <- c("F1", "F2", "F3", "F4", "F5", "F6")
rownames(X) <- c( "S1", "S2", "S3", "S4", "S5")
```


### Exercise 4

Print an element (3, 5) from matrix X:

```{r s3e4, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e4-hint}
# Use square brackets for matrix indexing

```

```{r s3e4-solution}
X[3, 5]
```

```{r s3e4-check}
test_1 <- find_statement("X\\[(.*)\\]", "you need to use square brackets", regex = TRUE)
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```

Print an second column from matrix X:

```{r s3e4i, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e4i-solution}
X[, 2]
```

```{r s3e4i-check}
test_1 <- find_statement("X\\[(.*)\\]", "you need to use square brackets", regex = TRUE)
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


Print an third row from matrix X:

```{r s3e4ii, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e4ii-solution}
X[3, ]
```

```{r s3e4ii-check}
test_1 <- find_statement("X\\[(.*)\\]", "you need to use square brackets", regex = TRUE)
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```

Print a submatrix of rows 1,2,3 and columns 4, 5 of matrix X:

```{r s3e4iii, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e4iii-solution}
X[1:3, 4:5]
```

```{r s3e4iii-check}
test_1 <- find_statement("X\\[(.*)\\]", "you need to use square brackets", regex = TRUE)
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```

### Exercise 5

Change elements (2, 5) matrix X to 0.

```{r s3e5, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e5-hint}
# Use an assignment operator and square brackets for matrix indexing 
```

```{r s3e5-solution}
X[2, 5] <- 0
```

```{r s3e5-check}
test_1 <- find_statement("X[2, 5] <- ", "you need to reassign 'X[2, 5]'")
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


Change all elements of the 1st column of matrix X to 2.

```{r s3e5i, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e5i-hint}
# Use an assignment operator and square brackets for matrix indexing 
```

```{r s3e5i-solution}
X[, 1] <- 2
```


```{r s3e5i-check}
test_1 <- find_statement("X[, 1] <- ", "you need to reassign 'X[, 1]'")
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```


Change the third row of matrix X to c(1,2,3,4,5,6).

```{r s3e5ii, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e5ii-hint}
# Use an assignment operator and square brackets for matrix indexing 
```


```{r s3e5ii-solution}
X[3, ] <- c(1,2,3,4,5,6)
```


```{r s3e5ii-check}
test_1 <- find_statement("X[3, ] <- ", "you need to reassign 'X[3, ]'")
USER_CODE %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, same_num(res)) 
```

### Exercise 6

Plot the F1 vs F2 columns of matrix X against each other.

```{r s3e6, exercise=TRUE,  exercise.setup = "pS3iii"}

```


```{r s3e6-solution}
plot(X)
```


Plot the F4 vs F5 columns of matrix X against each other.

```{r s3e6i, exercise=TRUE,  exercise.setup = "pS3iii"}

```

```{r s3e6i-solution}
plot(X[, 4:5])
```

The end of the section.

