---
title: "Programming with R"
tutorial:
  id: "com.exercises.programming"
  version: 1.0
output: 
  learnr::tutorial:
     progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = checkr::checkr_tutor)
```


## Logical Statements

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*


### Exercise 1

In this exercise will use data from a [reporsitory](https://github.com/tategallery/collection)
containing information around 70,000 artworks owned by [TATE](http://www.tate.org.uk/)
art galleries. Read the 'artist_data.csv' file (from https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv) 
into R:

```{r pS1}
artists_df <- read.csv("https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv")
```

```{r s1e1, exercise=TRUE}
artists_df <- 
```


```{r s1e1-hint}
# Use read.csv() function and supply the given URL address
```


```{r s1e1-solution}
artists_df <- read.csv("https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv")
```


```{r s1e1-check}
test_0 <- find_assignment("artists_df", "you need to create a data.frame named 'artists_df'")
test_1 <- find_call("read.csv(whatever)", "you need to use read.csv() function")
test_2 <- find_statement("https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv",
                         "wrong url address")
USER_CODE %>% test_0 %>% test_1 %>% test_2 
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```

###

Check the dimension of 'artists_df':

```{r s1e1i, exercise=TRUE, exercise.setup = "pS1"}

```

```{r s1e1i-solution}
dim(artists_df)
```

Inspect the structure of 'artists_df' using `str()` function:

```{r s1e1ii, exercise=TRUE, exercise.setup = "pS1"}

```

```{r s1e1ii-solution}
str(artists_df)
```

### 

Using square brackets, and `which()` function, extract a subset of artists who 
were born in the 19th century, i.e. those artists whose year of birth is from 
1801 to 1900 (inclusive):

```{r s1e1iii, exercise=TRUE, exercise.setup = "pS1"}
artists19th <- ...
```


```{r s1e1iii-hint}
# Use square brackets to subset the 'artists_df' and which() function 
# as well as '>=' and '<=' operators to select the artist of interest
```


```{r s1e1iii-solution}
artists19th <- artists_df[which(artists_df$yearOfBirth >= 1801 & artists_df$yearOfBirth <= 1900), ]
```


```{r s1e1iii-check}
test_0 <- find_assignment("artists19th", "you need to create a data.frame named 'artists19th'")
test_1 <- find_statement("(.*)\\[(.*)\\]", "you need to use square brackets", regex = TRUE)
USER_CODE %>% test_0 %>% test_1  
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```

###

```{r pS1i}
artists_df <- read.csv("https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv")
artists19th <- artists_df[artists_df$yearOfBirth >= 1801 & artists_df$yearOfBirth <= 1900, ]
```

What is the fraction of artists in 'artists_df' born in the 19th century?

```{r s1e1iv, exercise=TRUE, exercise.setup = "pS1i"}

```


```{r s1e1iv-hint}
# Find the number of rows in 'artists19th' and 'artists_df' to answer this question
```


```{r s1e1iv-solution}
nrow(artists19th)/nrow(artists_df)
# or 
dim(artists19th)[1]/dim(artists_df)[1]
```


```{r s1e1iv-check}
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```


###

Observe, that in the above example the function `which()` returns a set of indices 
where the condition `artists_df$yearOfBirth >= 1801 & artists_df$yearOfBirth <= 1900` 
evaluates to `TRUE` and allows us to select the rows with artist from the 19th century.
However, R also supports logical indexing and you can ommit the `which()` function:


```{r s1e1v, exercise=TRUE, exercise.setup = "pS1"}
artists19th <- artists_df[artists_df$yearOfBirth >= 1801 & artists_df$yearOfBirth <= 1900, ]
head(artists19th)
dim(artists19th)
```

**Note** that the above produces exactly the data.frame 'artists19th' results as before.

###

Now from the set of 19th century artists select the ones whose place of
birth was either London or Paris:

```{r s1e1vi, exercise=TRUE, exercise.setup = "pS1i"}
artists19th_lon_par <- 
```


```{r s1e1vi-hint}
# Again, use the square brackets notation but this time the '%in%' operator
# to select artists of interest
```


```{r s1e1vi-solution}
artists19th_lon_par <- artists19th[
  artists19th$placeOfBirth %in% c("London, United Kingdom", "Paris, France"), ]
```


```{r s1e1vi-check}
test_0 <- find_statement("artists19th", "you need to subset a data.frame named 'artists19th'")
test_1 <- find_statement("placeOfBirth", "you need to use 'find_statement' column to subset the data")
USER_CODE %>% test_0 %>% test_1  
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```


How many artists born in  the 19th century London and Paris were there in the
TATE collection?

```{r pS1ii}
artists_df <- read.csv("https://raw.githubusercontent.com/tategallery/collection/master/artist_data.csv")
artists19th <- artists_df[artists_df$yearOfBirth >= 1801 & artists_df$yearOfBirth <= 1900, ]
artists19th_lon_par <- artists19th[
  artists19th$placeOfBirth %in% c("London, United Kingdom", "Paris, France"), ]
```

```{r s1e1vii, exercise=TRUE, exercise.setup = "pS1ii"}

```

```{r s1e1vii-solution}
nrow(artists19th_lon_par)
```


```{r s1e1vii-check}
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```


###

How many artists in 'artists_df' have a missing year of death data?

```{r s1e1viii, exercise=TRUE, exercise.setup = "pS1"}

```


```{r s1e1viii-hint-1}
# To know which artists have missing year of death use is.na() function
```

```{r s1e1viii-hint-2}
# To count the artists use sum() pr nrow() if you previously subset the data
```


```{r s1e1viii-solution}
nrow(artists_df[is.na(artists_df$yearOfDeath), ])
# or
sum(is.na(artists_df$yearOfDeath))
```


```{r s1e1viii-check}
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "incorrect results")}, 
                substitute(res)))
```


## If/else statements

### Exercise 1

In this exercise, we will learn a bit about the game of [Blackjack](http://www.wikihow.com/Play-Blackjack).
Blackjack is a game where each player competes with the delealer,
and not other players. Each player attempts to beat the dealer by getting a count 
as close to 21 as possible, without exceeding it, ("busting"). Face cards (King, 
Queen, Jack) are each counted as 10, Ace can be counted as either 1 or 11, and 
any other card are counted as the numeric value shown on the card.

In each round of the game:

* At the beginning every player gets 2 face-up cards from the dealer; the dealer 
also receives 2 cards, but only one of them is face-up.
* After that, on their turn players may "hit" (take a card), "stand" (end their 
turn), "double" (double wager, take a single card and finish), "split" (if the 
two cards have the same value, separate them to make two hands) or "surrender" 
(give up a half-bet and retire from the game). Players continue
until they stand or go "bust", i.e. the total value of their hand exceeds 21.
* When the dealer has served every player, his face-down card is turned up. 
If the total is 17 or more, he must stand, otherwise he must 
continue to take cards until the total is 17 or more. 
If the dealer has an ace, and counting it as 11 would bring his 
total to 17 or more, he must stand. The dealer can't double, split or surrender.
* Lastly each player's hand is compared with the dealer's one and who comes 
closer to 21 becomes a winner.

More details on the game can be found at: 
http://www.bicyclecards.com/how-to-play/blackjack/

### Game setup

```{r pS2}
# Deck of cards
card_deck <- data.frame(card = rep(c(2:10, c("J", "Q", "K", "A")), 4),
                        value = rep(c(2:10, c(10, 10, 10, Inf)), 4),
                        stringsAsFactors = FALSE)
# Here we repeat indexing 6 times to stack 6 decks of cards
blackjack_cards <- card_deck[rep(1:nrow(card_deck), 6), ]

# Shuffle cards
set.seed(1234)
blackjack_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), size = nrow(blackjack_cards)), ]

# We rename the rows of 'blackjack_cards' to be the ordering of cards after the shuffle:
rownames(blackjack_cards) <- 1:nrow(blackjack_cards)

# Initial deal:
H <- blackjack_cards[1:2, ]
D <- blackjack_cards[3:4, ]
Drev <- D[1, ]
remain_cards <- blackjack_cards[5:nrow(blackjack_cards), ]

if (all(H$card != "A")) {
  V <- sum(H$value)
} else {
  H1 <- H[H$card != "A", ]
  no_aces<- nrow(H) - nrow(H1)
  # Note that only 1 of the aces can be counted as 11 otherwise you "bust"
  V1 <- sum(H1$value) + 11 + (no_aces- 1) 
  V2 <- sum(H1$value) + no_aces
  if (V1 > 21) { 
    # if counting ace as 11 result in "busting", count it as 1
    V <- V2
  } else {
    V <- V1 
  }
}

# Function for counting points in a hand 'H'
count_points <- function(H) {
  if (all(H$card != "A")) {
    V <- sum(H$value)
  } else {
    # Count points for non-aces:
    V0 <- sum(H$value[H$card != "A"])
    no_aces <- sum(H$card == "A")
    # Count one ace as 11
    V1 <- V0 + 11 + (no_aces - 1) 
    # Count all aces as 1
    V2 <- V0 + no_aces
    V <- ifelse(V1 > 21, V2, V1) 
  }
  return(V)
}


# Function for updating a hand
update_players_hand <- function(H, Drev, blackjack_cards) {
  V <- count_points(H)
  next_card <- blackjack_cards[1, ]
  if (V == 11 || all(c(V %in% c(9, 10), Drev$value < 7))) {
    A <- "double"
  } else if (V < 12) {
    A <- "hit"
  } else if (all(c(V >= 12, V <= 16)) && Drev$value > 7) {
    A <- "hit"
  } else if ("A" %in% H$card && all(c(V >= 13, V <= 17))) {
    A <- "hit"
  } else {
    A <- "stand"
  }
  if (A %in% c("double", "hit")) {
    H <- rbind(H, next_card)
    blackjack_cards <- blackjack_cards[-1, ]
  }
  return(list(A = A, H = H, remaining_cards = blackjack_cards))
}

# Function for initial deal
init_deal <- function(blackjack_cards, shuffle = TRUE) {
  if (shuffle == TRUE) {
    blackjack_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), nrow(blackjack_cards)), ]
  }
  H <- blackjack_cards[1:2, ]
  D <- blackjack_cards[3:4, ]
  # Remove players and dealers hand (4 crads) from the stack: 
  blackjack_cards <- blackjack_cards[5:nrow(blackjack_cards), ]
  return(list(player_hand = H, dealer_hand = D, remaining_cards = blackjack_cards))
}


# Dealer's turn
dealers_play <- function(D, remain_cards) {
  # Initial points:
  dealV0 <- sum(D$value[D$card != "A"])
  no_aces <- sum(D$card == "A")
  if (no_aces > 0) {
   dealV0 <- dealV0 + 11 + (no_aces - 1)
  }
  tot <- dealV0
  while (tot < 17) {
    next_card <- remain_cards[1, ]
    remain_cards <- remain_cards[-1, ]
    D <- rbind(D, next_card)
    if (next_card$card != "A") {
      tot <- tot + next_card$value
    } else if (tot + 11 > 17) {
      tot <- tot + 11
    } else {
      tot <- tot + 1
    }
  }
  return(list(D = D, remain_cards = remain_cards))
}


# players's turn
participants_play <- function(H, Drev, remain_cards) {
  stop <- FALSE
  while (!stop) {
    cur_turn <- update_players_hand(H, Drev, remain_cards)
    H <- cur_turn$H
    remain_cards <- cur_turn$remaining_cards
    V <- count_points(H)
    if (cur_turn$A %in% c("stand", "double") || V >= 21) {
      stop <- TRUE
    }
  }
  return(list(H = H, remain_cards = remain_cards))
}


blackjack_round <- function(blackjack_cards, shuffle = TRUE){
  # Set up the initial deal:
  init <- init_deal(blackjack_cards, shuffle)
  Drev <- init$dealer_hand[1, ]
  players_turn <- participants_play(init$player_hand, Drev, init$remaining_cards)
  dealers_turn <- dealers_play(init$dealer_hand, players_turn$remain_cards)
  remain_cards <- dealers_turn$remain_cards
  
  players_hand <- players_turn$H
  players_points <- count_points(players_turn$H)
  dealers_hand <- dealers_turn$D
  dealers_points <- count_points(dealers_turn$D)

  cat("Player's final hand:", players_hand$card, "\n")
  cat("Players's final points:", players_points, "\n")
  cat("Dealer's final hand:", dealers_hand$card, "\n")
  cat("Dealer's final points:", dealers_points, "\n")
  
  player_diff <- 21 - players_points
  dealer_diff <- 21 - dealers_points
  
  score <- c()
  if ((dealer_diff == player_diff) || (dealer_diff < 0 && player_diff < 0)) {
    cat("Player push with the dealer.")
    score <- c(score, 0)
  } else if (dealer_diff >= 0 && (dealer_diff < player_diff || player_diff < 0)) {
    cat("Player lost.")
    score <- c(score, -1)
  } else {
    cat("Player won!")
    score <- c(score, 1)
  }
  return(list(remain_cards = remain_cards, players_hand = players_hand, 
              players_points = players_points, dealers_hand = dealers_hand,
              dealers_points = dealers_points, score = score))
}
```

Here we build a data.frame 'card_deck' representing a standard deck 
of 52 cards with two columns: one denoting the card identity and its 
corresponding Blackjack value. Since, an ace in Blackjack 
does not have an univocal count, we will encode its value as 'Inf' (infinity) 
which will make it easy to detect and hand as a special case:

```{r s2e0, exercise = TRUE}
# Deck of cards
card_deck <- data.frame(card = rep(c(2:10, c("J", "Q", "K", "A")), 4),
                        value = rep(c(2:10, c(10, 10, 10, Inf)), 4),
                        stringsAsFactors = FALSE)
dim(card_deck)
# Here 'rep(x, n)' function repeats x, n number of times
```

At the casinos Blackjacks are usually played with 6 shuffled decks:

```{r s2e00, exercise = TRUE, exercise.setup = "pS2"}
set.seed(1234)
# Here we repeat indexing 6 times to stack 6 decks of cards
blackjack_cards <- card_deck[rep(1:nrow(card_deck), 6), ]

# Check that the number of rows in 'blackjack_cards' is 6x52:
dim(blackjack_cards)

# Shuffle cards
blackjack_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), size = nrow(blackjack_cards)), ]

# We rename the rows of 'blackjack_cards' to be the ordering of cards after the shuffle:
rownames(blackjack_cards) <- 1:nrow(blackjack_cards)

# Now we look at 10 first cards
knitr::kable(head(blackjack_cards, 10))
```

The cards are placed in a [dealer's shoe](https://en.wikipedia.org/wiki/Shoe_(cards)https://en.wikipedia.org/wiki/Shoe_(cards)).
and dealt consecutively.

### 

Imagine you are **the only player at the table** competing with the dealer.
Two cards first cards from teh 'blackjack_cards' are dealt to you. 
Write code contating**if/else statements** that compute the value, 'V', of your hand, 'H'. 
If the hand contains an ace, "A", count it as either 1 or 11, whichever 
achieves the value of the hand closer to 21, but not exceeding it.

```{r s2e1, exercise=TRUE, exercise.setup = "pS2"}
# Complete (...) in the following code:
H <- blackjack_cards[1:2, ]
if (...) {
  V <- ...
} else {
  ...
}
cat("The hand, 'H' = ", H$card, ", has value", V)
```


```{r s2e1-hint-1}
# Cases were the 'H' hand contains or does not contain aces
# should be addressed as separate if/else cnditions
```

```{r s2e1-hint-2}
# You can use 'all()' or 'any()' function to combine locgical statements
# Access the documentation using:
?all
?any
```


```{r s2e1-solution}
# This is an example solution
H <- blackjack_cards[1:2, ]

if (all(H$card != "A")) {
  V <- sum(H$value)
} else {
  H1 <- H[H$card != "A", ]
  no_aces<- nrow(H) - nrow(H1)
  # Note that only 1 of the aces can be counted as 11 otherwise you "bust"
  V1 <- sum(H1$value) + 11 + (no_aces- 1) 
  V2 <- sum(H1$value) + no_aces
  if (V1 > 21) { 
    # if counting ace as 11 result in "busting", count it as 1
    V <- V2
  } else {
    V <- V1 
  }
}
cat("The hand, 'H' = ", H$card, ", has value", V)
```

### 

After all players are dealt two cards, the dealer reveals one of his cards.

```{r s2e000, exercise=TRUE, exercise.setup = "pS2"}
D <- blackjack_cards[3:4, ]
Drev <- D[1, ]
Drev
```

###

Suppose, you are a novice in Blackjack and you play a [simplified basic strategy](http://blackjackcalculation.com/blackjack-super-easy-basic-strat.html),
i.e. you:

* Always double when your hand is 11
* Double 9 or 10 hand when the dealer has 2-6
* Hit when your is 12-16 when the dealer has 7-Ace
* Hit when your hand contains an ace and the value of the hand is 13-17
* Otherwise stand. 

Here we never 'split' to make the following exercises less complicated.


Write if/else statements to print your action, 'A', ("split", "double", "hit", "stand") 
for a hand 'H' constisting of two first cards in the stack with the value 'V' against 
the dealer's revealed card 'Drev' from above, using on the above strategy:   


```{r s2e1i, exercise=TRUE, exercise.setup = "pS2"}
# Complete (...) in the following code:
# You can use 'H', 'V' and 'Drev' variables created before, as they are stored in memory
if (...) {
  A <- "double"
} else if (...) {
  A <- "hit"
} else {
  A <- "stand"
}

cat("This turn, player", A)
```


```{r s2e1i-hint}
# You need to encode the given strategies as conditionals based
# on 'H' and 'Drev', use '&&', '||' ('and', 'or' operators) or
# preferred 'all()', 'any()' functions to combine logical statements.
# Function 'identical()' might also be useful if you want to compare
# vectors.
```


```{r s2e1i-solution}
# This is an example solution:

if (V == 11 || all(c(V %in% c(9, 10), Drev$value < 7))) {
  A <- "double"
} else if (V < 12) {
  A <- "hit"
} else if (all(c(V >= 12, V <= 16)) && Drev$value > 7) {
  A <- "hit"
} else if ("A" %in% H$card && all(c(V >= 13, V <= 17))) {
  A <- "hit"
} else {
  A <- "stand"
}

cat("This turn, player", A)
```

## Functions

### Exercise 1

Write a function which counts the points of a given Blackjack hand, 'H'
based on the code written in the previous section:

```{r s3e0, exercise = TRUE, exercise.setup = "pS2"}
H <- blackjack_cards[1:2, ]

if (all(H$card != "A")) {
  V <- sum(H$value)
} else {
  H1 <- H[H$card != "A", ]
  no_aces<- nrow(H) - nrow(H1)
  # Note that only 1 of the aces can be counted as 11 otherwise you "bust"
  V1 <- sum(H1$value) + 11 + (no_aces- 1) 
  V2 <- sum(H1$value) + no_aces
  if (V1 > 21) { 
    # if counting ace as 11 result in "busting", count it as 1
    V <- V2
  } else {
    V <- V1 
  }
}
cat("The hand, 'H' = ", H$card, ", has value", V)
```


The function should take a data.frame 'H' with two columns 'card' and 'value'
representing the current hand, and return its total point count.

```{r s3e1, exercise=TRUE, exercise.setup = "pS2"}
# Complete (...) in the following code:
count_points <- function(H) {
  ...
  V <- ...
  return(V)
}
```


```{r s3e1-hint}
# Use the code with if/else statements provided above
```


```{r s3e1-solution}
# This is an example solution
count_points <- function(H) {
  if (all(H$card != "A")) {
    V <- sum(H$value)
  } else {
    # Count points for non-aces:
    V0 <- sum(H$value[H$card != "A"])
    no_aces <- sum(H$card == "A")
    # Count one ace as 11
    V1 <- V0 + 11 + (no_aces - 1) 
    # Count all aces as 1
    V2 <- V0 + no_aces
    V <- ifelse(V1 > 21, V2, V1) 
  }
  return(V)
}
```


### Exercise 2

Suppose, you are playing th basic strategy as the one described
in the previous section. Write a function that takes your hand, 'H',
the dealer's revealed card, 'Drev', and the current 'blackjack_cards' 
stack as inputs and returns a lists consisting of player's action taken
and his or hers updated hand. 

'H', 'Drev' and 'blackjack_cards' are each a data.frame with columns 
'card' and 'value'. The function should return the value 
of the current hand 'H' then take an action corresponding to
the basic strategy as well as the remaining stack of cards:


```{r s3e2, exercise=TRUE, exercise.setup = "pS2"}
# Complete (...) in the following code:
update_players_hand <- function(H, Drev, blackjack_cards) {
  ...
  return(list(A = ..., H = ..., remaining_cards = ...))
}
# Here A is a string specifying an action and H is a data.frame representing the updated hand
```


```{r s3e2-hint}
#  To help you we provide some helper code which you should incorporate in your function:

V <- count_points(H)
next_card <- blackjack_cards[1, ]
if (V == 11 || all(c(V %in% c(9, 10), Drev$value < 7))) {
  A <- "double"
} else if (V < 12) {
  A <- "hit"
} else if (all(c(V >= 12, V <= 16)) && Drev$value > 7) {
  A <- "hit"
} else if ("A" %in% H$card && all(c(V >= 13, V <= 17))) {
  A <- "hit"
} else {
  A <- "stand"
}
if (A %in% c("double", "hit")) {
  H <- rbind(H, next_card)
}

```


```{r s3e2-solution}
# This is an example solution
update_players_hand <- function(H, Drev, blackjack_cards) {
  V <- count_points(H)
  next_card <- blackjack_cards[1, ]
  if (V == 11 || all(c(V %in% c(9, 10), Drev$value < 7))) {
    A <- "double"
  } else if (V < 12) {
    A <- "hit"
  } else if (all(c(V >= 12, V <= 16)) && Drev$value > 7) {
    A <- "hit"
  } else if ("A" %in% H$card && all(c(V >= 13, V <= 17))) {
    A <- "hit"
  } else {
    A <- "stand"
  }
  if (A %in% c("double", "hit")) {
    H <- rbind(H, next_card)
  }
  return(list(A = A, H = H, remaining_cards = blackjack_cards))
}
```

### Exercise 3

Write code setup an initial deal for a single player
and the dealer. If indicated, the function should first shuffle 
the cards. Then the function should assign a hand to the player, 'H', 
and the dealer, 'D' from the top of the card stack. The function should 
return a list with named elements and 'H', 'D' and 'blackjack_cards' 
corresponding to the player's hand, the dealer's hand, and the remaining 
stack if cards respectively. 

```{r s3e3, exercise=TRUE, exercise.setup = "pS2"}

init_deal <- function(blackjack_cards, shuffle = TRUE) {
  # Shuffle cards:
  ...
  # Player's hand:
  H <- ...
  # Dealer's hand:
  D <- ...
  # Remove player's and dealer's hands (4 crads) from the stack: 
  blackjack_cards <- ...
  return(list(player_hand = H, dealer_hand = D, remaining_cards = blackjack_cards))
}

```


```{r s3e3-hint-1}
# Here 'shuffle' is an optional argument, by defult set to TRUE.
# You should use an if statement shuffle the cards if 'shuffle' is set to TRUE. 
# Use 'sample()' to generate a permutation of  cards' indices and use it for shuffling.
```

```{r s3e3-hint-2}
# Extract the first 2 cards from the shuffled card stack using indexing 
# as the player's hand. The next two cards should be assignedc to the dealer.
```


```{r s3e3-solution}
# Example solution:
init_deal <- function(blackjack_cards, shuffle = TRUE) {
  if (shuffle == TRUE) {
    blackjack_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), nrow(blackjack_cards)), ]
  }
  H <- blackjack_cards[1:2, ]
  D <- blackjack_cards[3:4, ]
  # Remove players and dealers hand (4 crads) from the stack: 
  blackjack_cards <- blackjack_cards[5:nrow(blackjack_cards), ]
  return(list(player_hand = H, dealer_hand = D, remaining_cards = blackjack_cards))
}
```


### Exercise 4 (function calling)

Write code to simulate an initial deal for a sigle player and a dealer;
then prints the action taken by the player and his updated hand,
given that he or she follows the strategy provided.

```{r s3e4, exercise=TRUE, exercise.setup = "pS2"}
# Complete (...) in the following code:

# 'blackjack_cards' is saved in memory

# Set up the initial deal
init <- ...

# Use 'update_players_hand()' function created before 
# to get player's action and the updated hand:

...

cat("Player's action is:", ...)
cat("Player's hand is now: H =", ...)
```


```{r s3e4-solution}
# Set up the initial deal
init <- init_deal(blackjack_cards)

# Extract the player's hand
H <- init$player_hand

# Extract the dealer's revealed (first) card:
Drev <- init$dealer_hand[1, ]

# Remaining cards after the initial deal:
remain_cards <- init$remaining_cards

# Player's first action:
first_turn <- update_players_hand(H, Drev, remain_cards)

cat("Player's action is:", first_turn$A)
cat("Player's hand is now: H =", first_turn$H$card)
```



## Loops

### Exercise 1

In Blackjack, a round completes for a player whenever he or she 
decides to "stand" or goes "bust" (when the value of the player's 
hand exceeds 21). Write a `while` loop to simulate a complete round for 
a single player against the dealer. Print out te player's final hand 
and the corresponding value.


```{r s4e1, exercise=TRUE, exercise.setup = "pS2"}
# Set up the initial deal:
init <- init_deal(blackjack_cards)
H <- init$player_hand
Drev <- init$dealer_hand[1, ]
remain_cards <- init$remaining_cards

# Complete (...) in the following code:
stop <- FALSE
while (!stop) {
  # keep updating the player's hand until he doubles/stands/busts
  ...

}
cat("Dealer's revealed card:", Drev$card)
cat("Player's final hand is: H =", ...)
cat("Player's final points are: V =", ...)
```


```{r s4e1-solution}
# Set up the initial deal:
init <- init_deal(blackjack_cards)
H <- init$player_hand
Drev <- init$dealer_hand[1, ]
remain_cards <- init$remaining_cards

# Solution:
stop <- FALSE
while (!stop) {
  cur_turn <- update_players_hand(H, Drev, remain_cards)
  H <- cur_turn$H
  remain_cards <- cur_turn$remaining_cards
  V <- count_points(H)
  
  # Note that when doubling a player takes in one card and then "stands" 
  if (cur_turn$A %in% c("stand", "double") || V >= 21) {
    stop <- TRUE
  }
}
cat("Dealer's revealed card:", Drev$card)
cat("Player's final hand is: H =", H$card)
cat("Player's final points are: V =", V)
```

### 

Convert the code you just wrote to a function
which takes 'H' and 'remain_cards' as inputs
and returns a list with the players's final hand 
and the remaining stack:


```{r s4e1i, exercise=TRUE, exercise.setup = "pS2"}
participants_play <- function(H, remain_cards) {
  ...
}
```


```{r s4e1i-solution}
participants_play <- function(H, Drev, remain_cards) {
  stop <- FALSE
  while (!stop) {
    cur_turn <- update_players_hand(H, Drev, remain_cards)
    H <- cur_turn$H
    remain_cards <- cur_turn$remaining_cards
    V <- count_points(H)
    
    # Note that when doubling a player takes in one card and then "stands" 
    if (cur_turn$A %in% c("stand", "double") || V >= 21) {
      stop <- TRUE
    }
  }
  return(list(H = H, remain_cards = remain_cards))
}

```


### Exercise 2

 > When the dealer has served every player, his face-down card is turned up. 
 If the total is 17 or more, he must stand. If the total is 16 or under, 
 he must take a card. He must continue to take cards until the total is 17 
 or more, at which point the dealer must stand. If the dealer has an ace, 
 and counting it as 11 would bring his total to 17 or more (but not over 21), 
 he must count the ace as 11 and stand. The dealer's decisions, then, are 
 automatic on all plays, whereas the player always has the option of 
 taking one or more cards. - See more at: 
 http://www.bicyclecards.com/how-to-play/blackjack/#sthash.DaUmFTgL.dpuf

Write code simulating a dealer's turn using a `while` loop.
Assume 'D' the initial pair is loaded in memory.

```{r s4e2, exercise=TRUE, exercise.setup = "pS2"}
# Assume the remaining stack is stored in 'remain_cards' variable
# and the the dealer's initial deal is stored in 'D'

# Count the points for the initial pair:
dealV0 <- sum(D$value[D$card != "A"])
no_aces <- sum(D$card == "A")
if (no_aces > 0) {
  # Count one of the aces as 11
 dealV0 <- dealV0 + 11 + (no_aces - 1)
}

# Complete (...) in the following code:
dealV <- dealV0
while (...) {
  # Keep hitting until dealer's total >= 17
  ...
}

cat("Dealer's final hand:", ...)
cat("Dealer's final points:", ...)
```


```{r s4e2-solution}
# Count the points for the initial pair:
dealV0 <- sum(D$value[D$card != "A"])
no_aces <- sum(D$card == "A")
if (no_aces > 0) {
  # Count one of the aces as 11
 dealV0 <- dealV0 + 11 + (no_aces - 1)
}

tot <- dealV0
while (tot < 17) {
  # Dealer hits:
  next_card <- remain_cards[1, ]
  remain_cards <- remain_cards[-1, ]
  D <- rbind(D, next_card)
  if (next_card$card != "A") {
    tot <- tot + next_card$value
  } else if (tot + 11 > 17) {
    tot <- tot + 11
  } else {
    tot <- tot + 1
  }
}

cat("Dealer's final hand:", D$card)
cat("Dealer's final points:", count_points(D))
```


### 

Convert the code you just wrote to a function
which takes 'D' and 'remain_cards' as inputs
and returns the dealer's final hand:


```{r s4e2i, exercise=TRUE, exercise.setup = "pS2"}
dealers_play <- function(D, remain_cards) {
  ...
}
```


```{r s4e2i-solution}
dealers_play <- function(D, remain_cards) {
  # Initial points:
  dealV0 <- sum(D$value[D$card != "A"])
  no_aces <- sum(D$card == "A")
  if (no_aces > 0) {
   dealV0 <- dealV0 + 11 + (no_aces - 1)
  }
  tot <- dealV0
  while (tot < 17) {
    next_card <- remain_cards[1, ]
    remain_cards <- remain_cards[-1, ]
    D <- rbind(D, next_card)
    if (next_card$card != "A") {
      tot <- tot + next_card$value
    } else if (tot + 11 > 17) {
      tot <- tot + 11
    } else {
      tot <- tot + 1
    }
  }
  return(list(D = D, remain_cards = remain_cards))
}

```



### Exercise 3

Combining all the steps, one round of Blackjack looks like this:

```{r s4e0, exercise=TRUE, exercise.setup = "pS2"}
blackjack_round <- function(blackjack_cards, shuffle = TRUE){
  # Set up the initial deal:
  init <- init_deal(blackjack_cards, shuffle)
  Drev <- init$dealer_hand[1, ]
  players_turn <- participants_play(init$player_hand, Drev, init$remaining_cards)
  dealers_turn <- dealers_play(init$dealer_hand, players_turn$remain_cards)
  remain_cards <- dealers_turn$remain_cards
  
  players_hand <- players_turn$H
  players_points <- count_points(players_turn$H)
  dealers_hand <- dealers_turn$D
  dealers_points <- count_points(dealers_turn$D)

  cat("Player's final hand:", players_hand$card, "\n")
  cat("Players's final points:", players_points, "\n")
  cat("Dealer's final hand:", dealers_hand$card, "\n")
  cat("Dealer's final points:", dealers_points, "\n")
  
  player_diff <- 21 - players_points
  dealer_diff <- 21 - dealers_points
  
  score <- c()
  if ((dealer_diff == player_diff) || (dealer_diff < 0 && player_diff < 0)) {
    cat("Player push with the dealer.")
    score <- c(score, 0)
  } else if (dealer_diff >= 0 && (dealer_diff < player_diff || player_diff < 0)) {
    cat("Player lost.")
    score <- c(score, -1)
  } else {
    cat("Player won!")
    score <- c(score, 1)
  }
  return(list(remain_cards = remain_cards, players_hand = players_hand, 
              players_points = players_points, dealers_hand = dealers_hand,
              dealers_points = dealers_points, score = score))
}
```

###

Write a forloop to simulate 10 rounds of Blackjack
remeber to remove cards after every round. 
Shuffle only on first round. For every round
print out the player's and dealer's hand and points
as well as whether the player won or lost: 

```{r s4e3, exercise=TRUE, exercise.setup = "pS2"}
# Complete the following code
remain_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), nrow(blackjack_cards)), ]
for (round in 1:10) {
  
  ...
  
}
```

```{r s4e3-solution}
set.seed(1234)
remain_cards <- blackjack_cards[sample(1:nrow(blackjack_cards), nrow(blackjack_cards)), ]
for (r in 1:10) {
  cat("\n Round:", r, "\n")
  bj_round <- blackjack_round(remain_cards, shuffle = FALSE)
  remain_cards <- bj_round$remain_cards
}
```


### Exercise 4

Use `sapply()` function to generate a vector whose elements are equal to the total number of 
points for all non-ace cards (not "A"s) included in a stack composed of 1,2, .., 10 standard 
52-card-decks.

```{r s4e4, exercise=TRUE, exercise.setup = "pS2"}
sapply(1:10, function(i) {

})
```

```{r s4e4-solution}
sapply(1:10, function(i) {
  stack <- card_deck[rep(1:nrow(card_deck), i), ]
  tot <- sum(stack$value[stack$card != "A"])
})
```

The end of the tutorial.
