---
title: "Base Plotting"
author: Lan Huong Nguyen
tutorial:
  id: "com.exercises.base-plotting"
  version: 100.0
output: 
  learnr::tutorial:
     progressive: true
runtime: shiny_prerendered
---

options(tutorial.exercise.timelimit = 10)


```{r setup, include=FALSE}
library(learnr)
library(checkr)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = checkr::checkr_tutor)
```


## 

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*


### Exercise 1

**Download 'Credit.csv' file** from http://www-bcf.usc.edu/~gareth/ISL/data.html 
to your 'Downloads' directory. Set the 'current directory' to be the 'Downloads' folder,
which now contains the 'Credit.csv' file. 

```{r s1e1, exercise=TRUE}

```

```{r s1e1-hint}
# Use the following setwd() function
```

```{r s1e1-solution}
# Set the current directory to 'Downloads' folder folder:
setwd("~/Downloads/")
```


###

```{r, pDir}
setwd("~/Downloads/")
```


Use `file.show()` function to display the contants of the file 'Credit.csv'
in 'dataset' directory:

```{r s1e1i, exercise=TRUE, exercise.setup = "pDir"}

```

```{r s1e1i-solution}
file.show("Credit.csv")
```
