---
title: "Data into R"
author: Lan Huong Nguyen
tutorial:
  id: "com.exercises.data-to-R"
output: 
  learnr::tutorial:
     progressive: true
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(checkr)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = checkr::checkr_tutor)
tutorial_options(exercise.timelimit = 30)
```


## Importing data

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*


### Exercise 1

**Download 'Credit.csv' file** from http://www-bcf.usc.edu/~gareth/ISL/data.html 
to the 'Downloads' directory:

```{r s1e0, exercise=TRUE}
# Download to e.g. the Download directory
download.file("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv",
              destfile = "~/Downloads/Credit.csv")
```

Use `file.show()` function to display the contants of the file 'Credit.csv'
in 'Downloads' directory:


```{r s1e1i, exercise=TRUE, exercise.setup = "pDir"}

```

```{r s1e1i-solution}
file.show("~/Downloads/Credit.csv")
```


###

Now import the data into R using the `read.csv()` function
and assign it to 'credit' variable:

```{r s1e1ii, exercise=TRUE, exercise.setup = "pDir"}
credit <- 
```

```{r s1e1ii-solution}
credit <- read.csv("~/Downloads/Credit.csv")
# or 
# credit <- read.csv("Credit.csv")
```

```{r s1e1ii-check}
test_0 <- find_statement("credit", "you need to create a data.frame named 'credit'")
test_1 <- find_call("read.csv(whatever)", "you need to use `read.csv()` function")

USER_CODE %>% test_0 %>% test_1 
```


###

```{r pS1}
credit <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv")
```


Check what is the class of 'credit' variable

```{r s1e1iii, exercise=TRUE, exercise.setup = "pS1"}

```

```{r s1e1iii-solution}
class(credit)
```

```{r s1e1iii-check}
test_1 <- find_call("class(whatever)", "you need to use `class()` function")
test_2 <- find_statement("credit", "you need to use a data.frame 'credit'")

USER_CODE %>% test_1 %>% test_2
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "not correct results")}, 
                substitute(res)))
```

###

Print first six lines of 'credit'.
To nicely display the table in this html based tutorial.

```{r s1e1iv, exercise=TRUE, exercise.setup = "pS1"}

```

```{r s1e1iv-solution}
head(credit)
```

```{r s1e1iv-check}
test_1 <- find_statement("head", "you need to use `head()` function")
test_2 <- find_statement("credit", "you need to use a data.frame 'credit'")
USER_CODE %>% test_1 %>% test_2
```

###

Note that the first column contains the copy of the rownames.
This is because the original data in the "Credits.csv" file contains
a column for rownames. In R data.frames already have row names 
assigned separately, so you can remove this column.

Read the data into R again, but this time tell
R which column contains the row names. Check the `read.csv()` documentation
using `?read.csv`.



```{r s1e1v, exercise=TRUE, exercise.setup = "pDir"}

```


```{r s1e1v-hint}
# Use the row.names arguments in read.csv() function.
```

```{r s1e1v-solution}
credit <- read.csv("~/Downloads/Credit.csv", row.names = 1)
```

```{r s1e1v-check}
test_1 <- find_call("read.csv(whatever)", "you need to use `head()` function")
test_2 <- find_statement("row.names", "you should use a row.names argument")
USER_CODE %>% test_1 %>% test_2
```


###

```{r pS1i}
credit <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv", row.names = 1)
```

Use `head()` to again print first few lines of credit. Notice the change:

```{r s1e1vi, exercise=TRUE, exercise.setup = "pS1i"}

```


```{r s1e1vi-solution}
head(credit)
```


```{r s1e1vi-check}
test_1 <- find_statement("head", "you need to use `head()` function")
test_2 <- find_statement("credit", "you need to use a data.frame 'credit'")
USER_CODE %>% test_1 %>% test_2
```

**Note** that now there is no column with the row names.

### 

The data in 'Credits.csv' is used for examples in *An Introduction to Statistical Learning* book 
and the file was downloaded from http://www-bcf.usc.edu/~gareth/ISL/Credit.csv 

It is very convenient feature of the R read function including `read.csv()`
and `read.table()` that data hosted online can be directly read into R
without a prior download to your local machine. See the following:


```{r s1e1vii, exercise=TRUE}
credit <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv", row.names = 1)
head(credit)
```

Observe that 'credit' data.frame is the same as before.


### Exercise 2

**Download** 'College.csv' file fromhttp://www-bcf.usc.edu/~gareth/ISL/data.html 
to 'Downloads' folder:

```{r s1e00, exercise=TRUE}
# Download to e.g. the Download directory
download.file("http://www-bcf.usc.edu/~gareth/ISL/College.csv",
              destfile = "~/Downloads/College.csv")
```


Then use `scan()` function to read the first 10 lines
of the data:

```{r s1e2,  exercise=TRUE, exercise.setup = "pDir"}

```

```{r s1e2-hint-1}
# Use function scan() and specify 'what' to be "" or character()
```

```{r s1e2-hint-2}
# Specify the 'sep' argument to be "\n" so that every line is read separately as
# one element of a list.
```

```{r s1e2-hint-3}
# Specify the 'nlines' argument to be 10.
```

```{r s1e2-solution}
scan("~/Downloads/College.csv", "", sep = "\n", nlines = 10)
```



```{r s1e2-check}
test_1 <- find_call("scan(whatever)", "you need to use `scan()` function")
test_2 <- find_call("scan(whatever, whatever)", "you need to specify 'what' argument")
test_3 <- find_statement("College.csv", "you need toread 'Colleges.csv' file")
test_4 <- find_statement("sep", "you need to specify a 'sep' argument")
test_5 <- find_statement("nlines", "you need to specify a 'nlines' argument")

USER_CODE %>% test_1 %>% test_2 %>% test_3 %>% test_4 %>% test_5
```


### Exercise 3

NASA's online data portal includes many interesting datasets you can download.
One of them is a datasets listing all the NASA facilities available
at http://open.nasa.gov/datasets/NASA_Labs_Facilities.xlsx.

Download the 'NASA_Labs_Facilities.xlsx' file to your 'Downloads' directory:

```{r s1e3_1, exercise=TRUE}
url1 <- "http://open.nasa.gov/datasets/NASA_Labs_Facilities.xlsx"

# Download to e.g. the Download directory
download.file(url1, destfile = "~/Downloads/NASA_Facilities.xlsx")
```


If you havent already, please uncomment and install package 'readxl' which 
has functionalities for reading Excel. Then, load the package into memory:

```{r s1e3_0, exercise=TRUE}
# install.packages("readxl", repos="http://cran.us.r-project.org")
library(readxl)
```

```{r p_s1e3}
library(readxl)
nasa_fc <- read_excel("~/Downloads/NASA_Facilities.xlsx")
```

###

Read the file into 'nasa_fc' variable using `read_excel()` function.
You can check the function documentation by typing `?read_excel`.

```{r s1e3, exercise=TRUE, exercise.setup = "p_s1e3"}
nasa_fc <- 
```


```{r s1e3-solution}
nasa_fc <- read_excel("~/Downloads/NASA_Facilities.xlsx")
```

```{r s1e3-check}
test_0 <- find_assignment("nasa_fc", "you need to read into 'nasa_fc' variable")
test_1 <- find_call("read_excel(whatever)", "you need to use `read_excel()` function")
USER_CODE %>% test_0 %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "not correct results")}, 
                substitute(res)))
```


###


```{r s1e3_2, exercise=TRUE, exercise.setup = "p_s1e3"}
class(nasa_fc)
```

`read_excel()` output is of class 'tibble'. Convert it to a more familiar 
data.frame class:

```{r s1e3i, exercise=TRUE, exercise.setup = "p_s1e3"}

```

```{r s1e3i-solution}
nasa_fc.df <- as.data.frame(nasa_fc)
```

```{r s1e3i-check}
test_0 <- find_statement("nasa_fc", "you need to read into 'nasa_fc' variable")
test_1 <- find_call("as.data.frame(whatever)", "you need to use `as.data.frame()` function")
USER_CODE %>% test_0 %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "not correct results")}, 
                substitute(res)))
```


```{r p_s1e3i}
nasa_fc <- readxl::read_excel("~/Downloads/NASA_Facilities.xlsx")
nasa_fc.df <- as.data.frame(nasa_fc)
```

The class of 'nasa_fc.df' is now of class:

```{r s1e3_3, exercise=TRUE, exercise.setup = "p_s1e3i"}
class(nasa_fc.df)
```


### 

Print the first 10 lines of 'nasa_fc.df':

```{r s1e3ii, exercise=TRUE, exercise.setup = "p_s1e3i"}

```


```{r s1e3ii-solution}
head(nasa_fc.df, 10)
```

```{r s1e3ii-check}
test_0 <- find_call("head(whatever)", "you need to use `head()` function")
test_1 <- find_call("head(whatever, whatever)", "you need to use two arguments of `head()`.")

USER_CODE %>% test_0 %>% test_1 
soln_test(USER_CODE, SOLN_CODE, res = final_, 
          same_(function(S, R, hint = hint) {ifelse(identical(S, R), "", "not correct results")}, 
                substitute(res)))
```

### 

**Note** that R generated some column names e.g. "X__1", this is because
the Excel file we used included some merged columns. If you'd like to remove
the top merged column names, use the 'skip' argument when calling 'read_excel()' function:

```{r s1e3iii, exercise = TRUE, exercise.setup = "p_s1e3"}
nasa_fc <- read_excel("~/Downloads/NASA_Facilities.xlsx", skip = 1)
nasa_fc.df <- as.data.frame(nasa_fc)
head(nasa_fc.df)
```

```{r "p_s1e3iv"}
nasa_fc <- readxl::read_excel("~/Downloads/NASA_Facilities.xlsx", skip = 1)
nasa_fc.df <- as.data.frame(nasa_fc)
```

### 

Note that the first column denotes the row index, but is incorrectly
labelled 'Agency'. Change the name of the first column to 'ID',
and the second column one to 'Agency' and print first few lines of
the altered data.frame:

```{r s1e3iv, exercise = TRUE, exercise.setup = "p_s1e3iv"}

```


```{r s1e3iv-solution}
colnames(nasa_fc.df)[1] <- "ID"
colnames(nasa_fc.df)[2] <- "Agency"
head(nasa_fc.df)
```


The end of the section.

## Saving data

*Enter your code directly into the chunks and press the 'Run Code' button to execute it.
There are multiple exercises for each topic, you must complete an exercise and press 'Continue'
to move onto the next one. Only once a tick mark appears next to every exercise should you 
be able to continue to the next section.*


### Exercise 1

Set the current directory to be the 'Documents' folder.

```{r s2e00, exercise=TRUE}
setwd("~/Documents/")
```


###

Suppose you have created a following data.frame storing information
on some characters from Harry Potter:


```{r s2e0, exercise=TRUE}
hp_info <- data.frame(first = c("Harry", "Ron", "Hermione", "Ginny"), 
                      last = c("Potter", "Wesley", "Granger", "Wesley"),
                      gender = c("M", "M", "F", "F"),
                      date = c(31, 1, 19, 11),
                      month = c("Jul","March", "Sep", "Aug"))
```


###

```{r, pS2}
setwd("~/Documents/")
hp_info <- data.frame(first = c("Harry", "Ron", "Hermione", "Ginny"), 
                      last = c("Potter", "Wesley", "Granger", "Wesley"),
                      gender = c("M", "M", "F", "F"),
                      date = c(31, 1, 19, 11),
                      month = c("Jul","March", "Sep", "Aug"))

nasa_fc <- readxl::read_excel("~/Downloads/NASA_Facilities.xlsx", skip = 1)
nasa_fc.df <- as.data.frame(nasa_fc)
colnames(nasa_fc.df)[1] <- "ID"
colnames(nasa_fc.df)[2] <- "Agency"

credit <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv ", row.names = 1)
```

Write (save) the 'hp_info' data.frame to a file 'harry_potter.txt'

```{r s2e1i, exercise = TRUE, exercise.setup = "pS2"}

```

```{r s2e1i-solution}
write.table(hp_info, file="harry_potter.txt")
```

###

```{r s2q1}
question('Where was the file saved?',
    answer('in Desktop'),
    answer('in Documents', correct = TRUE),
    answer('in the current directory', correct = TRUE),
    answer('in a temporary directory'),
    answer('in Downloads'),
    random_answer_order = TRUE,
    allow_retry = TRUE
  )
```


###

Save 'nasa_fc.df' and 'credit' data generated earlier in the session 
to file 'mydata.rdata':

```{r s2e1iii, exercise=TRUE, exercise.setup = "pS2"}

```

```{r s2e1iii-hint}
# Use save() function
```

```{r s2e1iii-solution}
save(nasa_fc.df, credit, file="mydata.rdata")
```

###

Use `ls()` function to print out all the variables/objects generated in the session
and saved inmemory:

```{r, s2e11, exercise = TRUE, exercise.setup = "pS2"}
ls()
```

You might have noticed an unfamiliar *server_context* variable.
This object was not generated by you, but was created by R for the
purpose of running this tutorial application. 


###

Remove all data generated in the session using `rm()` function:

```{r, s2e111, exercise = TRUE}
rm(list = ls())
```


Check that now memory is empty:

```{r pS22}
rm(list = ls())
```


```{r, s2e1111, exercise = TRUE, exercise.setup = "pS22"}
ls()
```



### 

Load all the objects saved in 'mydata.rdata' and check that 'nasa_fc.df' and 
'credit' objects are restored:

```{r, s2e1vi, exercise = TRUE}

```

```{r, s2e1vi-hint}
# Use load() to restore the data and and ls() to print out all the variables
```

```{r, s2e1vi-solution}
load("mydata.rdata")
ls()
```


The end of the tutorial.
